Activity Prediction with Random Forest Algorythm
========================================================

In this article we show how to get predictions about personal activity from a device that records such activities. 

We first load datasets into data frames:

```{r}
trainingData <- read.csv("pml-training.csv")
testData <- read.csv("pml-testing.csv")
```

Then we take a look at the head of the data for an insight on which data woudl be useful for our purposes.

```{r}
head(trainingData, n=3)
```

We see that many of the variables have either missing data or NA for data. We exclude thsoe and make a list of useful data variables:



```{r}
useful_variables <- c("roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt",
                      "gyros_belt_x", "gyros_belt_y", "gyros_belt_z",
                      "accel_belt_x", "accel_belt_y", "accel_belt_z", 
                      "magnet_belt_x", "magnet_belt_y",
                        "magnet_belt_z", "roll_arm", "pitch_arm",
                        "yaw_arm", "total_accel_arm", "gyros_arm_x",
                        "gyros_arm_y", "gyros_arm_z", "accel_arm_x",
                      "accel_arm_y", "accel_arm_z", "magnet_arm_x", "magnet_arm_y", 
                        "magnet_arm_z", "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell",
                      "gyros_dumbbell_y", "gyros_dumbbell_z", "accel_dumbbell_x",
                      "accel_dumbbell_y", "accel_dumbbell_z", "magnet_dumbbell_x",
                        "magnet_dumbbell_y", "magnet_dumbbell_z",
                      "roll_forearm", "pitch_forearm", "yaw_forearm",
                      "gyros_forearm_x", "gyros_forearm_y",
                      "gyros_forearm_z", "accel_forearm_x", 
                      "accel_forearm_y", "accel_forearm_z", "magnet_forearm_x",
                      "magnet_forearm_y", "magnet_forearm_z", "classe"
                      )
```

We create the new "clean" data frames with only the complete sets of observations

```{r}
cleanTrainingData <- trainingData[useful_variables]
```

Then we create the splitdf function that will return a list of training and testing sets

```{r}
splitdf <- function(dataframe, seed=NULL) {
        if (!is.null(seed)) set.seed(seed)
        index <- 1:nrow(dataframe)
        trainindex <- sample(index, trunc(length(index)/2))
        trainset <- dataframe[trainindex, ]
        testset <- dataframe[-trainindex, ]
        list(trainset=trainset,testset=testset)
}
```

Next, we apply the function on the cleaned training data in order to get a new training and test sets:

```{r}
splits <- splitdf(cleanTrainingData, seed=808)
```

Wesave the training and testing sets as data frames

```{r}
training <- splits$trainset
testing <- splits$testset
```

For our predictive algoryth we will use the Random Forest algorythm. First, we load the Random Forest library

```{r}
library(randomForest)
```

We fit the randomForest model

```{r}
model <- randomForest(classe~., 
                      data = training, 
                      importance=TRUE,
                      keep.forest=TRUE
)
```

```{r}
print(model)
```

We see that the error rate for this model and this training data is 0.82%. We expect the error rate with the test set to be little larger, about 1%.

Next, we take a look at the importance of various variables for this model.

```{r}
varImpPlot(model, type=1)
```


Next, we use the model to make the predictions on the cleaned test Data set:
```{r}
predicted <- predict(model, newdata=testing[ ,-50])
```


We want to compare the predicted data to the actual data. For that purpose we look at the number of miscategorized data points
```{r}
actual <- testing$classe
sum(actual != predicted)
```

We see that there are 98 incorrectly predicted observations. Since out testing data has a total of 9811 observations, this error rate is almsot exactly 1%, as we had expected.

Finally, we take a look at the testData. We clean it, and then we make the prediction on it based on our model.

```{r}
useful_variables_2 <- useful_variables[1:49]
cleanTestData <- testData[useful_variables_2]
predicted_2 <- predict(model, cleanTestData)
predicted_2
```

We eneter these predictions into the course testing site. All of the predictions turn out to be correct. This validates our modeling algorythm. 

